\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-08-26 paper.qmd
Updated Figure 17

error_df <- error_df |>
  mutate(type = case_when(
    row_wise_abs_error <= 2 ~ "error 0-2",
    row_wise_abs_error <= 4 ~ "error 2-4",
    row_wise_abs_error <= 6 ~ "error 4-6",
    row_wise_abs_error <= 6 ~ "error 6-8",
    row_wise_abs_error <= 10 ~ "error 8-10",
    .default = "error greter than 10"
  )) |>
  mutate(type = factor(type, levels = c(
    "error 0-2", "error 2-4", "error 4-6", "error 6-8", "error 8-10", 
    "error greter than 10")))

@@

```{r}
#| echo: false
#| fig-cap: "$10\\text{-}D$ model error in PaCMAP layout of MNIST digit 1 dataset in $2\\text{-}D$. Most low $10\\text{-}D$ model errors are distributed along the lower edge of the $2\\text{-}D$ structure, while most large $10\\text{-}D$ model errors are concentrated along the upper edge (a). There are certain data points that exhibit high error rates due to their deviation from the usual $10\\text{-}D$ data structure, which makes them anomalies. These anomalies can be classified into two types: those that are anomalies within the non-linear structure and those that lie outside of it. The images associated with high model error points within the non-linear structure display different patterns of the digit 1 (b). However, when comparing these images to the ones found outside of the non-linear structure, it becomes evident that the latter display different patterns of the digit 1 (c). "
#| label: fig-model-mnist
#| fig-pos: H

error_plot_mnist 
```

@@

::: {#fig-mnist-anomalies layout-ncol="2" fig-pos="H"}
![](figures/mnist/img_error_sample_within.png){#fig-mnist1-within}

![](figures/mnist/img_error_sample_out.png){#fig-mnist1-out}

Some images of handwritten digit 1 which occur high model error (a) within the non-linear structure, and (b) outside the non-linear structure. The images shows different patterns of digit 1.
:::

>>>>

error_df <- error_df |>
  mutate(type = case_when(
    row_wise_abs_error <= 2 ~ "error 0-2",
    row_wise_abs_error <= 4 ~ "error 2-4",
    row_wise_abs_error <= 6 ~ "error 4-6",
    row_wise_abs_error <= 8 ~ "error 6-8",
    row_wise_abs_error <= 10 ~ "error 8-10",
    .default = "error greter than 10"
  )) |>
  mutate(type = factor(type, levels = c(
    "error 0-2", "error 2-4", "error 4-6", "error 6-8", "error 8-10", 
    "error greter than 10")))

@@

```{r}
#| label: img-with-error

## Data with pixel values
mnist_data <- read_rds("data/mnist/mnist_digit_1.rds")

img_error_within <- c(6169, 5987, 773, 6671, 4785, 4699, 5173, 5475)

pixels_gathered_within <-  mnist_data |>
  mutate(instance = row_number()) |>
  gather(pixel, value, -Label, -instance) |>
  extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
  filter(instance %in% img_error_within)

imge_error_sample_within <- pixels_gathered_within |>
  ggplot(aes(x, y, fill = value)) +
  geom_tile() +
  facet_wrap(~ instance, ncol = 4) +
  coord_fixed() +
  scale_fill_continuous_sequential(palette = "Grays") +
  theme(strip.background = element_blank(),
        strip.text.x = element_blank(),
        legend.position = "none")

img_error_outside <- c(1489, 5095, 855, 3143, 453, 3619, 1015, 409) #, 5101

pixels_gathered_outside <-  mnist_data |>
  mutate(instance = row_number()) |>
  gather(pixel, value, -Label, -instance) |>
  extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
  filter(instance %in% img_error_outside)

imge_error_sample_outside <- pixels_gathered_outside |>
  ggplot(aes(x, y, fill = value)) +
  geom_tile() +
  facet_wrap(~ instance, ncol = 4) +
  coord_fixed() +
  scale_fill_continuous_sequential(palette = "Grays") +
  theme(strip.background = element_blank(),
        strip.text.x = element_blank(),
        legend.position = "none")

```{r}
#| label: fig-model-mnist
#| fig-cap: "$10\\text{-}D$ model error in PaCMAP layout of MNIST digit 1 dataset in $2\\text{-}D$. Most low $10\\text{-}D$ model errors are distributed along the lower edge of the $2\\text{-}D$ structure, while most large $10\\text{-}D$ model errors are concentrated along the upper edge (a). There are certain data points that exhibit large error rates due to their deviation from the usual $10\\text{-}D$ data structure, which makes them anomalies. These anomalies can be classified into two types: those that are anomalies within the non-linear structure and those that lie outside of it. The images associated with high model error points within the non-linear structure display different patterns of the digit 1 (b). However, when comparing these images to the ones found outside of the non-linear structure, it becomes evident that the latter display different patterns of the digit 1 (c). "
#| fig-pos: H
#| fig-width: 4
#| fig-height: 8
#| out-height: 80%

error_plot_mnist / (imge_error_sample_within + imge_error_sample_outside) +
  plot_annotation(tag_levels = 'a') &
  theme(plot.tag.position  = c(0.92, 0.92),
        plot.tag = element_text(color = "grey70"))
```

@@

<!-- ::: {#fig-mnist-anomalies layout-ncol="2" fig-pos="H"} -->
<!-- ![](figures/mnist/img_error_sample_within.png){#fig-mnist1-within} -->

<!-- ![](figures/mnist/img_error_sample_out.png){#fig-mnist1-out} -->

<!-- Some images of handwritten digit 1 which occur high model error (a) within the non-linear structure, and (b) outside the non-linear structure. The images shows different patterns of digit 1. -->
<!-- ::: -->

### 2024-08-26 paper.tex
Updated Figure 17

\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-model-mnist-1.pdf}

@@

\caption{\label{fig-model-mnist}\(10\text{-}D\) model error in PaCMAP
layout of MNIST digit 1 dataset in \(2\text{-}D\). Most low
\(10\text{-}D\) model errors are distributed along the lower edge of the
\(2\text{-}D\) structure, while most large \(10\text{-}D\) model errors
are concentrated along the upper edge (a). There are certain data points
that exhibit high error rates due to their deviation from the usual
\(10\text{-}D\) data structure, which makes them anomalies. These
anomalies can be classified into two types: those that are anomalies
within the non-linear structure and those that lie outside of it. The
images associated with high model error points within the non-linear
structure display different patterns of the digit 1 (b). However, when
comparing these images to the ones found outside of the non-linear
structure, it becomes evident that the latter display different patterns
of the digit 1 (c).}

\begin{minipage}{0.50\linewidth}

\includegraphics{figures/mnist/img_error_sample_within.png}

\subcaption{\label{fig-mnist1-within}}

\end{minipage}%
%
\begin{minipage}{0.50\linewidth}

\includegraphics{figures/mnist/img_error_sample_out.png}

\subcaption{\label{fig-mnist1-out}}

\end{minipage}%

\caption{\label{fig-mnist-anomalies}Some images of handwritten digit 1
which occur high model error (a) within the non-linear structure, and
(b) outside the non-linear structure. The images shows different
patterns of digit 1.}

>>>>

\includegraphics[width=1\textwidth,height=0.8\textheight]{paper_files/figure-pdf/fig-model-mnist-1.pdf}

@@

\caption{\label{fig-model-mnist}\(10\text{-}D\) model error in PaCMAP
layout of MNIST digit 1 dataset in \(2\text{-}D\). Most low
\(10\text{-}D\) model errors are distributed along the lower edge of the
\(2\text{-}D\) structure, while most large \(10\text{-}D\) model errors
are concentrated along the upper edge (a). There are certain data points
that exhibit large error rates due to their deviation from the usual
\(10\text{-}D\) data structure, which makes them anomalies. These
anomalies can be classified into two types: those that are anomalies
within the non-linear structure and those that lie outside of it. The
images associated with high model error points within the non-linear
structure display different patterns of the digit 1 (b). However, when
comparing these images to the ones found outside of the non-linear
structure, it becomes evident that the latter display different patterns
of the digit 1 (c).}
